---
title: "Hugo Academic Tutorial"
author: "Daniela Cassol"
output: html_document
---

# Annotations to create and edit template

## Install Package

```{r, eval=FALSE}
install.packages("blogdown")
blogdown::install_hugo()
```

## Create the new website template

```{r, eval=FALSE}
blogdown::new_site(theme = "gcushen/hugo-academic")
```

## Build and View the WebSite

```{r, eval=FALSE}
blogdown::build_site()
blogdown::serve_site()
```

## Delete Demo Section

- Delete files:
  - ```content/home/demo.md```
  - ```content/home/hero.md```

## Change Colors:
 - Check name in ```config/_default/params.toml```
    - theme = "minimal" ## Could change the theme here or create a new file and added the colors. 
    - ```themes/hugo-academic/data/themes``` ## Config files for the themes
    
## Edit About
  - ```content/authors/admin/_index.md```: edit
  - replace the image and keep the same name
  - ```content/home/about.md```: Title `Biography`

## Edit the menu

```config/_default/menus.toml```


## Edit `Contact` Section
  - Edit the info here: ```config/_default/params.toml```


# Themes

- Gallery of [Hugo Themes](https://themes.gohugo.io/)
  1. Select the theme
  2. Click on `Download` (it will be redirected to GitHub Repository)
  3. Replace the theme argument to the username/reponame, for example: 
      - ```blogdown::new_site(theme = "MeiK2333/github-style")```

# Reference

- [blogdown: Creating Websites with R Markdown](https://bookdown.org/yihui/blogdown/)

