---
title: "The WondeRs of R Markdown"
author: "Marília P. Gaiarsa"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    highlight: tango
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
thanks: R-Ladies Riverside
---

# Brief Introduction
R Markdown is a file format for making dynamic documents with R, It allows you to execute code and generate high quality documents. R Markdown documents are fully reproducible and support lots of different output formats (such as pdf, html, and doc).

It can be used as  "modern day lab notebook where you can capture not only what you did, but also what you were thinking". 

## Application Examples

1. Books
2. [Thesis/Dissertations](https://github.com/ismayc/thesisdown)
3. Websites
4. This document!

## File structure
**Header**: The text at the top of the document, written in YAML format surrounded by `---` .

**Markdown sections**: Text that describes your workflow written using markdown syntax.

**Code chunks**: Chunks of R code that can be run and also can be rendered using knitr to an output document.

## Creating an R Markdown File
File -> new file -> R Markdown (!)

It is going to ask you for a name, and the output wanted (you can change it later). Then you have a default example that we can play with =) 

When you first knit, it will ask you to save your file. 

Done! You can now start tweaking your document.


## Compiling an R Markdown File
There are a couple ways to compile an R Markdown file:

1. Press the `Knit` button (!)
2. Keyboard shortcut: `Ctrl + Shift + K (Cmd + Shift + K on macOS)`
3. In your console type: `rmarkdown::render()`

Note, R Markdown files **ONLY** use datasets, packages, objects, parameters, etc. defined or included in the file (not those defined in the R session or R environment). 

# Header (YAML metadata)

YAML metadata is the information you provide to format and generate an R Markdown file.

The YAML metadata composes the first few lines of the R Markdown file, and includes the `title`, `author`, `date` and `output` type.

As a starting point, the YAML data of your document can be:
```
---
title: "R-Ladies Meeting"
author: "R-Ladies"
date: "01/09/2020"
output: html_document
---
```

You can also add more details and refine the layout a bit:
```
---
title: "Doing It All With R Markdown"
author: "Marilia P. Gaiarsa"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: paper
    highlight: tango
---
```

Click [here](https://www.datadreaming.org/post/r-markdown-theme-gallery/) for a theme gallery. 

# Text in R Markdown

Text in R Markdown is written with Markdown syntax (!). As with (most) things in R, a [cheat sheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) is your best friend!

## Inline Formatting

* Italics: `_text_` or `*text*` renders _text_ and *text* respectively
* Bold: `**text**` produces **text**
* Verbatim: `` `Y~n` `` results in `Y~n`
* Subscript: `Y~n~` gives Y~n~
* Superscript: `X^2^` produces X^2^
* Inline Code: ` `code` ` 
* Hyperlinks: `[text](link)`
* Images: `![alt text or image title](path/to/image)`
* Footnotes: `^[This is a footnote]`

(BibTeX is recommended for citations)

## Block-level formatting

Section headers are written as:
```
# First-level Header

## Second-level Header

### Thid-level Header
```

Unordered list items start with `*`, `-`, or `+`, and they can be nested:
```
Consider this list: 

* item
* item
* item
    * sub-item
    * sub-item
* item
```
The output is:

* item
* item
* item
    * sub-item
    * sub-item
* item

Ordered list items start with numbers and letters:
```
Consider this list: 

1. item
2. item
3. item
    a. sub-item
    b. sub-item
4. item
```
The output is similar to the Ordered list:

Consider this list: 

1. item
2. item
3. item
    a. sub-item
    b. sub-item
4. item

*There must be an empty line rigth before the list, in order for R Markdown to notice your list is a list.*

## Equations and LaTex

You can use LaTeX equations' format (click [here](https://www.overleaf.com/learn/latex/Mathematical_expressions) for some examples of LaTex expressions). 

Inline equations can be wrapped in a pair of dollar signs, such as ```$f(x) = {n \choose k} p^{k} (1-p)^{n-k}$``` which results in $f(x) = {n \choose k} p^{k} (1-p)^{n-k}.$

Similarly, math expressions can be written in double dollar signs, such as ```$$f(x) = {n \choose k} p^{k} (1-p)^{n-k}$$``` which produces $$f(x) = {n \choose k} p^{k} (1-p)^{n-k}$$

# R Chunks 

An R code chunk can be inserted by using the Rstudio toolbar (the `Insert` button) or the keyboard shortcut `Ctrl + Alt + I (Cmd + Option + I on macOS)` (or by simply typing ` ``` `). 

Both will result in:
````
```{r}`r ''`

```
````

## R Chunk Options

R chunks can be used to produce text output, tables, or graphics.

Chunk options control all output types and can be provided inside the curly brackets: 
``` ```{r, chunk-label, results='hide', fig.height=4}```

Useful chunk options:

* ```eval```: whether to evaluate a code chunk
* ```echo```: whether to echo the source code in the output file
* ```results```: when set to ``` 'hide'```, text output will be hidden; when set to ``` 'asis'```, regular output is provided. 
* ```collapse```: whether to merge text output and source code into a single code block in the output. 
* ```warning```, ```message```, ```error```: whether to show warning, message, and errors in the output file. 
* ```include```: whether to include anything from the code chunk. 
* ```cache```: whether to enable caching, i.e, whether or not the chunk of code will be evaluated the next time the document is compiled. 
* ```fig.width``` and ```fig.height```: graphical size of R plots in inches. 
* ```out.width``` and ```out.height```: output size of R plots in the output file. e.g., ```out.width = '80%'``` means 80% of the page width.
*```fig.align```: alignment of plots, ```'left'```, ```'center'```, or ```'right'```. 

(click [here](https://yihui.org/knitr/options/) for more chunk options examples)

Now you try!

````
```{r}`r ''`
library(ggplot)
```
````


```{r}
library(ggplot2)
```

How to make it not print a warning?

````
```{r}`r ''`
library(ggplot)
```
````

```{r, warning=FALSE}
library(ggplot2)
```

## Other Chunk Types

Chunks are of many programming languages. You can specify a chunk to be in R, Python, Julia, C++, SQL, etc. 

The variety of languages engines come from the **knitr** package.

# Figures and Tables

We are going to use Teresa's code from our last meeting to plot some stuff! Thus, we will need to load the `ggplot` package and require the `iris` dataset. 

Remember to specify `include=FALSE` to ensure that the chunk does not appear in our output file but is still evaluated. 


````
```{r, include=FALSE}`r ''`
library('ggplot2')
data(iris)
```
````


```{r setup, include=FALSE}
library('ggplot2')
data(iris)
```

## Figures 

Then we can include any of the plots we created earlier by including the code that renders the plot:

````
```{r}`r ''`
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point()
```
````

```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point()
```

You can also tweak the plot a bit. Here is a horrible example:
````
```{r, fig.height=1, fig.width=10}`r ''`
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point()
```
````

```{r, fig.height=1, fig.width=10}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point()
```

## Tables
Let's use the iris dataset and wrangle it a bit (note in the code how we include `message=FALSE` and `warning=FALSE`):

```{r, message=FALSE, warning=FALSE}
library('ggplot2')
library('tidyverse')
nMeanSamples <- iris %>% 
  group_by(Species) %>% 
  summarise(mean.sep.l = mean(Sepal.Length),
            mean.sep.w = mean(Sepal.Width),
            n())
head(nMeanSamples)
```

The easiest way to include tables is to use ```knitr::kable()```:

````
```{r table, echo=FALSE}`r ''`
knitr::kable(nMeanSamples)
```
````

```{r table, echo=FALSE}
knitr::kable(nMeanSamples)
```

# Resources

The presentation and tutorial were adapted from [Maricela F. Cruz](https://github.com/rladies/meetup-presentations_irvine/blob/master/R-LadiesIrvine_ggplot2_rmarkdown_11142018/R%20Markdown%20Tutorial.Rmd) (R-Ladies Irvine) and from [Malcolm Barrett](https://github.com/laRusers/presentations/blob/master/2018-11-12_rmarkdown_hpc/r-markdown-intro/presentation.Rmd) (LA R Users).

1. [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/html-document.html)
2. [Other Cheat Sheets](https://www.rstudio.com/resources/cheatsheets/)
4. [Chunk Options](https://yihui.name/knitr/options.)
5. [https://rmarkdown.rstudio.com/](https://rmarkdown.rstudio.com/)
6. [https://rmarkdown.rstudio.com/lesson-1.html](https://rmarkdown.rstudio.com/lesson-1.html)
7. [https://rmarkdown.rstudio.com/gallery.html](https://rmarkdown.rstudio.com/gallery.html)
8. [Cool intro](https://ourcodingclub.github.io/2016/11/24/rmarkdown-1.html)