---
title: 'Analyzing Genomics Data in R with Bioconductor'
author: 'Stephanie Hicks'
date: '2020-07-22'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", out.width = '100%')
library(tidyverse)
library(cowplot)
library(lubridate)
library(here)
```

```{r, install-pkgs, eval=FALSE}
install.packages("BiocManager")
BiocManager::install("BiocPkgTools")
BiocManager::install("GenomicRanges")
```



## Explore Bioconductor statistics with `BiocPkgTools`

Functions to access metadata around Bioc packages and usage in a tidy data format. 

```{r}
suppressMessages({
  library(BiocPkgTools)
})

pkgs <- biocDownloadStats()
head(pkgs)
```

## How many Bioconductor packages in 2019? 

```{r, message=FALSE}
pkgs %>% 
  filter(Month != "all", Year==2020) %>%
  select(Package, repo) %>% 
  distinct() %>%
  group_by(repo) %>%
  count()
```

* Annotation packages = streamlines tedious bookkeeping 
* Experiment data packages = contains processed data; useful for teaching

## Plot of Bioconductor packages over time

```{r}
pkgs %>% 
  filter(Month != "all", repo=="Software") %>%
  select(Package, Year) %>%
  distinct() 
```

```{r}
pkgs %>% 
  filter(Month != "all", repo=="Software") %>%
  select(Package, Year) %>%
  distinct() %>%
  group_by(Year) %>%
  summarize(total_packages = n()) %>%
  arrange(desc(Year)) %>%
  ggplot(aes(x=Year,y=total_packages)) + 
    geom_point() + 
    xlab("Year") +
    ylab("Package count") +
    labs(title = "Number of Bioconductor Software Packages")
```

## Creating a `GenomicRanges`

```{r, message=FALSE}
suppressMessages({
  library(GenomicRanges)
})

gr <- GRanges(seqnames = "chr1", strand = c("+", "-"),
              ranges = IRanges(start = c(102012,520211),
                               end=c(120303, 526211)), 
              gene_id = c(1001,2151), 
              score = c(10, 25))
gr
```

```{r}
width(gr)
```

----
```{r}
gr[gr$score > 15, ]
```


----
```{r, warning=FALSE, message=TRUE}
suppressMessages({
  library(plyranges)
})

gr %>% 
  filter(score > 15)
```

----
```{r, warning=FALSE, message=TRUE}
gr %>% 
  filter(score > 15) %>% 
  width()
```

