---
title: "Semi-Tidy Maps"
author: "Tessa Shates"
date: "3/21/2021"
output:
  html_document: default
  pdf_document: default
---

#### Code so we can more easily run the show/pre-troubleshoot problems 
Created primarily by TS so far with contributions from TB, and refined/edited by DC & MG.  

### First, we must go through the libraries to download into R 
```{r, message = FALSE}
library(tidyverse)
library(lubridate) ## allows for parsing the character dates into a date format, and has functions to pull out parts of the date e.g. year

library(sf)
library(leaflet)
library(mapview)
library(huxtable)
library(gdtools)
library(systemfonts)
library(flextable)

library(ggplot2)
#devtools::install_github("dkahle/ggmap") #latest version of GGMAP
library(ggmap) #THIS IS OPTIONAL IF YOU WANT TO DO THE GOOGLE API KEY
library(ggsn)
library(dplyr)
library(tibble)
library(ggpubr)
library(vegan)

### FOR FUN COLORS :) 
library("devtools")
#devtools::install_github("katiejolly/nationalparkcolors") 
#if you want this you need devtools installed
#devtools::install_github("johannesbjork/LaCroixColoR")
library(LaCroixColoR)
library("nationalparkcolors") 

library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(rgeos)
```


Read in the CSV, reduce and evaluate the dataset.  

```{r, message = FALSE, results= FALSE}
## Read in the data set
ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv") 

## Learn about the data
head(ufo_sightings)

## Subset data to only one country
#let's subset to just USA (or just any country you want)
ufo2<- subset(ufo_sightings, country == "us")

## Omit NAs 
ufo3<- na.omit(ufo2)

## Remove "unknown" UFO shapes 
ufo4<- ufo3[ufo3$ufo_shape != "unknown",] 

## How large is our data set? 
dim(ufo4)
ncol(ufo4) #number of columns
nrow(ufo4) #number of rows 
```

#### Teresa's Data Wrangling with the dates  
The date format on the file is pretty clunky. 

```{r, message = FALSE, results= FALSE}
## Let's deal with the dates
head(ufo4$date_time) ## looks like we have Month day year hour minute format. (mdy_HM)
ufo5 <- ufo4 %>% ## read this like take dataset ufo4 THEN, do the next line to it
  mutate(
    date_time = parse_date_time(date_time, 'mdy_HM'), ## convert the string to a date for lubridate
    year = year(date_time),
    month = month(date_time) ##...we can pull out more info from this if you want, but you get the gist here. 
  ) ## mutate adds columns to your dataframe without changing any of the existing ones dimensions etc.

## Compare ufo4 & ufo5 datasets
head(ufo4)
head(ufo5) # has new columns year & month
```
### Define variables & do some base R plots   
In this section, I will also shrink to certain types of UFO shapes & focus on CA data points only 

```{r, message = FALSE, results= FALSE}
##BASE R SIMPLE MAP##
gx = ufo5$longitude
gy = ufo5$latitude

## Start with a basic plot. 
plot(gx, gy, xlab = "Latitude", ylab = "Longitude", main="USA UFO Sightings")

## This is quite a lot of points! Even without a background map, we can still see a vague USA outline. If you want to continue with this, best to further subset based on category of your interest. 

## Reducing to only one state in the dataset 
ufoCA<-ufo5[ufo5$state == "ca",] 

## Now let's get picky about which UFO shapes we will include
ufoCA$ufo_shape #to print out the different shape options

## I am reducing to fireballs, formations, flashes, flares and disks. 
ufo6<-ufoCA[ufoCA$ufo_shape %in% c("fireball","formation","flash","flare","disk"),]

## Base plot again to see how this differs
rx = ufo6$longitude
ry = ufo6$latitude

## Start with a basic plot. 
plot(rx, ry, xlab = "Latitude", ylab = "Longitude", main="CA UFO Sightings")

```


### GGPLOT TIME  
TBH I do not know what the alpha=0.4 is that TB added :) 

Simple ggplot, followed by TB summarize bubble plot by n_sightings. 

```{r, message = FALSE, results= FALSE}
## Simplest GGPLOT FOR CALIFORNIA
ufo6 %>% 
  ggplot()+
  geom_point(mapping = (aes(x = longitude, y = latitude)), alpha=0.4)+
  ggtitle("CA UFO Sightings")+
  theme_classic() 


## Fancier TB Bubble plots
## Let's summarize by sightings in any given city
cities_ufo <- ufo6 %>% 
  group_by(city_area) %>% 
  summarize(
    latitude=mean(latitude), ## because summarize is going to shrink our rows we want to keep this data in it. We could use it as a grouping variable, but likely there are lat long differences within city areas?
    longitude=mean(longitude),
    n_sightings = length(description)
  )

## You can do the same thing for any state (if you didn't shrink it to just CA you could summarize all the states- dataset ufo5), just sub state in for city area in the group_by line. You can also group by multiple factors, e.g. ufo shape within state.

## Beginnings of a bubble plot...
cities_ufo %>% 
  ggplot()+
  geom_point(mapping = (aes(x = longitude, y = latitude, size=n_sightings)), alpha=0.4)+
  ggtitle("CA UFO Sightings by city")+
  theme_classic() 
#so fancy oo love it
```

#### Now let us add the boundaries & backgrounds  

Start with the natural earth package and data set, then zoom into west coast. 

```{r, message = FALSE, results= FALSE}
## Start with just plotting the world 
# Pull out the dataset
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world) # What is type class of data? It is a dataframe. 

#plot the whole world - plots all country borders
ggplot(data = world) +
  geom_sf()

#plot again but with titles
#and a subtitle on the number of sovereign nations in the dataset
ggplot(data = world) +
  geom_sf() +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("World map", subtitle = paste0("(", length(unique(world$sovereignt)), " countries)"))

##Previously, I looked up some boundary lat long for west coast & California
ggplot(data = world) +
    geom_sf(color="black", fill="lightgreen") +
  coord_sf(xlim = c(-126.15, -111.12), ylim = c(31, 42), expand = FALSE)+
  ggtitle("USA West Coast")+
  theme_classic()


##For fun, let's use one of the colors from the LaCroix Color palette
##Previously, I looked up some boundary lat long for west coast & California
ggplot(data = world) +
    geom_sf(color="black", fill="#E9E4A6") +
  coord_sf(xlim = c(-126.15, -111.12), ylim = c(31, 42), expand = FALSE)+
  ggtitle("USA West Coast")+
  theme_classic()
```  


#### Now let us bring back the UFO data in here  
- Here I use the map made by the natural earth package and dataset, zoom into the west coast, and add geom_points from UFO sightings in California. In the first plot, I have the color of the point change for UFO shape.  
- I also bring back TB's bubbleplot, and put in on the yellow west coast map background.  

```{r, message = FALSE, results= FALSE}
##Same as the last plot, but adding geom_point 
#Automatically fills in different colors based on ufo_shape from the ggplot2 package. 
ggplot(data = world) +
    geom_sf(color="black", fill="#E9E4A6") +
  coord_sf(xlim = c(-126.15, -111.12), ylim = c(31, 42), expand = FALSE)+
  ggtitle("CA UFO Sightings")+
  theme_classic()+
geom_point(data=ufo6, mapping = (aes(x = longitude, y = latitude, col=ufo_shape)), size=1, stroke=1)

#Let's bring back Teresa's bubbleplot
  ggplot(data = world) +
    geom_sf(color="black", fill="#E9E4A6") +
  coord_sf(xlim = c(-126.15, -111.12), ylim = c(31, 42), expand = FALSE)+
  geom_point(data=cities_ufo, mapping = (aes(x = longitude, y = latitude, size=n_sightings)), alpha=0.4)+
  ggtitle("CA UFO Sightings by city")+
  theme_classic() 


```

#### Interactive maps with leaflet  

Make sure to use the smaller subset dataset, too much and R will struggle to deal with it.  
You can click on points to see the information about the data point.  
If you do this in the console, you can also easily export and **Save as image**. 
Learn more about this package and its capabilities here: https://rstudio.github.io/leaflet/ 

```{r, message = FALSE, results= FALSE}
mapview(ufo6, xcol = "longitude", ycol = "latitude", crs = 4269, grid = FALSE)
```

#### Maps with ggplot2 and ggmap (Google Earth)  

In order to do this yourself, you need to obtain a Google API Key.  
The first 1000 requests are free, and after it is quite cheap. No more than $3/month for most uses.  
1. [API key pricing](https://cloud.google.com/maps-platform/pricing)  
2. [CRAN-R instructions for API key & ggmpap usage](https://cran.r-project.org/web/packages/ggmap/readme/README.html)  


```{r, message = FALSE, results= FALSE, eval=FALSE}
#Map can be satellite, terrain, streetview, etc. 
#You can use a center point like below, or pick a city as a center
#If you aren't sure what lat, lon to use, you can create the map with 
#mapview ! 
#Least zoomed in is 1, most is 21 (building level) 
map1 <- get_googlemap(center = "Sacramento", zoom = 6, maptype = "terrain")
ggmap(map1) #to view the map with a center in Sacramento 
#or
map2 <- get_googlemap(center = c(lon = -120.60791, lat = 37.91603), zoom = 6, maptype = "satellite")
ggmap(map2)

## Here we use map2 to be a backdrop for our ufo sightings in california. 
ggmap(map2)+
  ggtitle("UFO Sightings in California") + 
  theme_void() + #this theme removes the latitude longitude lines and ticks marks. There are other themes on the ggplot cheatsheet that keep them. 
  geom_point(data = ufo6, mapping = (aes(x = longitude, y = latitude, col=ufo_shape)), size=1, stroke=1)
```


